---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```
<!-- badges: start -->
[![cran version](http://www.r-pkg.org/badges/version/rasterdiv)](https://cran.r-project.org/package=rasterdiv)
[![Downloads (monthly)](https://cranlogs.r-pkg.org/badges/last-month/rasterdiv?color=brightgreen)](https://cranlogs.r-pkg.org/badges/last-month/rasterdiv)
[![Downloads (total)](https://cranlogs.r-pkg.org/badges/grand-total/rasterdiv?color=brightgreen)](https://www.r-pkg.org/pkg/rasterdiv)

<!-- badges: end -->

# rasterdiv

## How to install?

### The stable version can be installed from the [CRAN](https://cran.microsoft.com/):

```{r, eval = FALSE}
install.packages("rasterdiv")
```

### The development version can be installed from [GitHub](https://github.com/):

```{r, eval = FALSE}
library(devtools)
remote::install_github("mattmar/rasterdiv")
```

## What is rasterdiv?

*__rasterdiv__* is an R package that provides functions to apply diversity indexes based on Information Theory on RasterLayer or numerical matrices. Supported indexes are:

- Parametric Rao’s quadratic entropy (classical and multidimensional);
- Shannon's diversity index;
- Pielou's evenness index;
- Hill's generalised entropy;
- Rényi's generalised entropy;
- Berger-Parker index;
- Cumulative Residual Entropy (CRE)

*__rasterdiv__* uses a "moving window" approach to derive such indexes; however, area-based (e.g., by using vector layers) Parametric Rao’s quadratic entropy is also possible through the [*paRao()*](https://mattmar.github.io/rasterdiv/articles/rasterdiv_area_based_Rao.html) function. Read more on *__rasterdiv__*:

- [*__rasterdiv__—An Information Theory tailored R package for measuring ecosystem heterogeneity from space: To the origin and back*](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.13583);
- [*Measuring diversity from space: a global view of the free and open source rasterdiv R package under a coding perspective*](https://link.springer.com/article/10.1007/s42974-021-00042-x);
- [*From zero to infinity: Minimum to maximum diversity of the planet by spatio-parametric Rao’s quadratic entropy*](https://onlinelibrary.wiley.com/doi/full/10.1111/geb.13270)
