---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dpi = 300,             # High resolution for sharp plots
  fig.width = 7,         # Width of the figures
  fig.height = 5,        # Height of the figures
  out.width = "100%",    # Adjust output width to be responsive
  out.height = "auto",   # Adjust output height to maintain aspect ratio
  fig.align = 'center'   # Center align figures
)
```

<!-- badges: start -->
[![cran version](http://www.r-pkg.org/badges/version/rasterdiv)](https://cran.r-project.org/package=rasterdiv)
[![Downloads (monthly)](https://cranlogs.r-pkg.org/badges/last-month/rasterdiv?color=brightgreen)](https://cranlogs.r-pkg.org/badges/last-month/rasterdiv)
[![Downloads (total)](https://cranlogs.r-pkg.org/badges/grand-total/rasterdiv?color=brightgreen)](https://www.r-pkg.org/pkg/rasterdiv)
<!-- badges: end -->

# rasterdiv

## Installation

### Stable version from [CRAN](https://cran.microsoft.com/):

```{r, eval = FALSE}
install.packages("rasterdiv")
```

### Development version from [GitHub](https://github.com/):

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("mattmar/rasterdiv")
```

## What is rasterdiv?

*__rasterdiv__* is an R package that provides functions to apply diversity indexes based on Information Theory on RasterLayer or numerical matrices. Supported indexes include:

- Parametric Rao’s quadratic entropy (classical and multidimensional);
- Shannon's diversity index;
- Pielou's evenness index;
- Hill's generalised entropy;
- Rényi's generalised entropy;
- Berger-Parker index;
- Cumulative Residual Entropy (CRE)


## New Features
* Time Series Integration: The package recently integrated time series of remote sensing data to derive Rao's index, for example considering plant phenology. See the vignette: [Multidimension Rao’s Index](https://mattmar.github.io/rasterdiv/articles/rasterdiv_03_Advanced_multidimension_Rao.html).
* Helical Graphs: Prepare and plot time series with “helical graphs”. See the vignette: `r if(knitr::is_html_output()) {
  "[Visualising rasterdiv indexes with Helical Plots](https://mattmar.github.io/rasterdiv/articles/rasterdiv_05_Helical_Plots.html)"
} else {
  "\\href{https://mattmar.github.io/rasterdiv/articles/rasterdiv_05_Helical_Plots.html}{Visualising rasterdiv indexes with Helical Plots}"
}`.

## Key Features
* Moving Window and Area-Based Calculations: Derive indexes using a “moving window” approach (all functions) or area-based Parametric Rao’s quadratic entropy via [*paRao()*](https://mattmar.github.io/rasterdiv/articles/rasterdiv_area_based_Rao.html)

## Learn More on **rasterdiv**
- [*__rasterdiv__—An Information Theory tailored R package for measuring ecosystem heterogeneity from space: To the origin and back*](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.13583);
- [*Measuring diversity from space: a global view of the free and open source rasterdiv R package under a coding perspective*](https://link.springer.com/article/10.1007/s42974-021-00042-x);
- [*From zero to infinity: Minimum to maximum diversity of the planet by spatio-parametric Rao’s quadratic entropy*](https://onlinelibrary.wiley.com/doi/full/10.1111/geb.13270)
- [*Helical graphs to visualize the NDVI temporal variation of forest vegetation in an open source space*](https://www.sciencedirect.com/science/article/abs/pii/S157495412200406X)
- [*Phenological Diversity Trends with Remote Sensing Datacubes*](https://www.researchgate.net/profile/Matteo-Marcantonio/publication/381288639_Phenological_Diversity_Trends_with_Remote_Sensing_Datacubes/links/6666a0fbb769e76919269fa1/Phenological-Diversity-Trends-with-Remote-Sensing-Datacubes.pdf)
