---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```
<!-- badges: start -->
<!-- [![cran version](http://www.r-pkg.org/badges/version/rasterdiv)](https://cran.r-project.org/package=rasterdiv) -->
<!-- [![Downloads (monthly)](https://cranlogs.r-pkg.org/badges/last-month/rasterdiv?color=brightgreen)](https://cranlogs.r-pkg.org/badges/last-month/rasterdiv)
 -->
<!--[![Downloads (total)](https://cranlogs.r-pkg.org/badges/grand-total/rasterdiv?color=brightgreen)](https://www.r-pkg.org/pkg/rasterdiv) -->
<!-- badges: end -->

# How to install?

### Stable versions can be installed from the CRAN:

```{r, eval = FALSE}
install.packages("rasterdiv")
```

### Development versions can be installed with:

```{r, eval = FALSE}
library(devtools)
install_github("mattmar/rasterdiv")
```

## What is rasterdiv?

__rasterdiv__ is a package for the R statistical software and environment. It aims to provide functions to apply Information Theory based diversity indexes on RasterLayer or numerical matrices, such as Shannon's entropy or Cumulative Residual Entropy (CRE).
