---
title: "Multidimension Rao's Index."
author: "Matteo Marcantonio, Daniele da Re, Duccio Rocchini"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Multidimension Rao's Index.}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE ,results='hide'}
knitr::opts_chunk$set(fig.width=12, fig.height=8, out.width = "95%", fig.align = "center", fig.dpi = 150, collapse = TRUE, comment = "#") 
```

```{r results='hide', message=FALSE, warning=FALSE}
library(terra)
library(gstat)
library(rasterdiv)
library(rasterVis)
library(viridis)
```

## Overview
In this vignette, we demonstrate how to compute the multidimensional Rao's Index using **rasterdiv** for multiple numerical matrices, represented as spatially autocorrelated *SpatRaster* objects.

# Creating Autocorrelated Spatial Patterns
First, we establish a grid that will be used to create spatial patterns through a semivariogram model.

```{r results='hide', message=FALSE, warning=FALSE}
gridDim <- 40
xy <- expand.grid(x=1:gridDim, y=1:gridDim)
```
Next, we define a semivariogram and use it to simulate autocorrelated spatial data.

```{r results='hide', message=FALSE, warning=FALSE}
varioMod <- vgm(psill=0.005, range=100, model='Exp')
zDummy <- gstat(formula=z~1, locations = ~x+y, dummy=TRUE, beta=200, model=varioMod, nmax=1)
set.seed(123)
xyz <- predict(zDummy, newdata=xy, nsim=2)
```

With the simulated data, we then create two SpatRaster objects that could represent environmental conditions like plant functional traits.

```{r results='hide', message=FALSE, warning=FALSE}
# Define coordinate reference system
utm32N <- "+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"
# Create SpatRasters
r <- terra::rast(nrow=40, ncol=40, crs=utm32N, ext=ext(0,10000, 0,10000))
r1 <- r
# Populate with simulated values
values(r) <- xyz$sim1
values(r1) <- xyz$sim2
```
## Computing Multidimensional Rao's Index

We now calculate the multidimensional Rao's Index using varying window sizes and alpha values.

```{r results='hide', message=FALSE, warning=FALSE, eval=FALSE}
mRao <- paRao(x=list(r, r1), window=c(3, 5), alpha=c(1, Inf), na.tolerance=1, method="multidimension", simplify=3)
```
The output is a nested list of SpatRaster which we can transform in a stack of SpatRaster and plotted together with the input layers, as follows:

## Visualisation of the result

```{r results='hide', message=FALSE, warning=FALSE, eval=FALSE}
# Create a list of all the rasters to plot
rasters_to_plot <- c(r, r1, mRao[[1]]$alpha.1, mRao[[2]]$alpha.Inf)
names(rasters_to_plot) <- c("Raster1", "Raster2", "Rao_Index_Window_3", "Rao_Index_Window_5")
```
```{r results='hide', message=FALSE, warning=FALSE, include = FALSE}
rasters_to_plot <- readRDS("rasters_to_plot.RDS")
```

```{r results='hide', message=FALSE, warning=FALSE}
# Use lapply to create a list of levelplots
plots_list <- lapply(rasters_to_plot, function(rst) {
  levelplot(rst, margin=FALSE, 
  col.regions=magma(100), 
  main=list(label=names(rst), 
  cex=1.5))
})

# Arrange the plots in a grid
do.call(gridExtra::grid.arrange, c(plots_list, ncol = 2))
```